---
title: "R Notebook"
output: html_notebook
---



# IN2019_v03 ==============================================================================
# Code associated with student profile in Marine Matters
# Figure 1: Map ==========================================================================

# Libraries ------------------------------------------------------------------------------

```{r}
library(rgdal)                # for spTransform() & project()
library(ncdf4)                # for current data and SST data downloaded from AODN
library(ggplot2)              # for ggplot()
library(ggrepel)              # for geom_text_repel() - repel overlapping text labels
library(data.table)
library(tidyverse)
library(oce)                  # for a colorset suitable for oceanography
library(rnaturalearth)        # for a map
```

# Variables ------------------------------------------------------------------------------


# Load data ------------------------------------------------------------------------------

# Map data

```{r}
world <- ne_countries(scale = "Medium", returnclass = "sf")


# Load environmental table
# The original source for this data is https://www.cmar.csiro.au/data/trawler/survey_details.cfm?survey=IN2019_V03

in19v03.env<-read_csv('~/Dropbox/voyages/IN2019_V03/in2019_v03HydroDep.csv')
```

# option to load NRS data

## CURRENT DATA ##
# ======================================================================================
# Create a simple localised map of currents based on data from IMOS Ocean Current using ggplot.
# For this example we selected the 'Gridded sea level anomaly - Delayed mode' product,
# set a bounding box for the Australian region (lat, lon) and date range covering the duration
# of the voyage (IN2019_v03) and requested a download in .nc format. This  section of code is  
# based on a script sourced from the IMOS user code library (https://github.com/aodn/imos-user-code-library/blob/master/R/demos/acorn.R)
# 
#  Parameters for this subset
#  -----------------------------
#  Collection: IMOS - OceanCurrent - Gridded sea level anomaly - Delayed mode
#  Spatial: From Lat/Lon -60.0,90.0 to Lat/Lon 90.0,180.0
#  Temporal: 2019-May-25-00:00-UTC to 2019-Jun-15-00:00-UTC
# 
#  source: https://portal.aodn.org.au
# ======================================================================================
# Load current data

```{r}
curr<-nc_open("~/Downloads/IMOS_aggregation_20200402T214601Z/IMOS_aggregation_20200402T214601Z.nc");
lon <- ncvar_get(curr, 'LONGITUDE')
lat <- ncvar_get(curr, 'LATITUDE')
vcur <- ncvar_get(curr, 'VCUR')
ucur <- ncvar_get(curr, 'UCUR')

## Transform lon and lat objects so that they have the same length as udata and vdata

long <- rep(lon, length(lat))
lati <- NA
for (i in 1:length(lat)){
latitude <- rep(lat[i], length(lon))
if (i == 1) lati <- latitude else lati <- c(lati, latitude)
}
lon_b <- lon; lon <- long
lat_b <- lat; lat <- lati

vdata<-vcur[,,10]  #northward geostrophic velocity
udata<-ucur[,,10]  #eastward geostrophic velocity
dat1 <- list( )
dat1$x <- c( lon)
dat1$y <- c( lat)

x1 <- c( lon) + c( udata)/1 # Divide u values by 5 to resize the length of the current directional arrow
y1 <- c( lat) + c( vdata)/1 # Divide v values by 5 to resize the length of the current directional arrow

## Subsample the current vectors to obtain an appropriate spacing for the scale of the figure

dat2<-dat1
sub<-seq(1,length(dat1$x),3)

dat2$x <- dat1$x[sub]
dat2$y <- dat1$y[sub]
x2<-x1[sub]
y2<-y1[sub]

# ggplot() + geom_point(data=in19v03.env, aes(x=Longitude, y=Latitide), color='black')

## set your lat and lon ranges here 
xrange <- c(100,125)
yrange <- c (-42,-8)
```
## Plot a basic map with deployments and currents
```{r}
ggplot() + 
  geom_point(data=in19v03.env, aes(x=Longitude, y=Latitude), color='black') +
  geom_segment(data=as.data.frame(dat2), aes(x=x, xend=x2, y=y, yend=y2), lwd=0.1, arrow = arrow(type = 'open',length = unit(0.05, "cm"))) + 
  coord_fixed(1) + 
  geom_sf(data = world, lwd = 0) + coord_sf(xlim = c(xrange), ylim=c(yrange))+
  theme_bw() 


## Adding labels is relatively straight forward with geom_label_repel


ggplot() + 
  geom_point(data=in19v03.env, aes(x=Longitude, y=Latitude), color='black') +
  geom_segment(data=as.data.frame(dat2), aes(x=x, xend=x2, y=y, yend=y2), lwd=0.1, arrow = arrow(type = 'open',length = unit(0.05, "cm"))) + 
  geom_label_repel(nudge_x=-5,segment.alpha = 0.4,label.size = 0.1, data=in19v03.env %>% 
          distinct(Deployment, .keep_all =T) %>% 
          filter (round(Latitude,0) %%2==0),aes(y=Latitude, x=Longitude, label=Deployment)) +
  coord_fixed(1) + 
  geom_sf(data = world, lwd = 0) + coord_sf(xlim = c(xrange), ylim=c(yrange))+
  theme_bw() 
```


## SST DATA ##
# ======================================================================================
# Create a simple localised map of Sea Surface Temperature based on data from IMOS Ocean Color using ggplot.
# For this example we selected the 'SRS - SST - L3S - Single Sensor - 6 day - day time - Australia' product,
# set a bounding box for the Australian region (lat, lon) and date range covering the duration
# of the voyage (IN2016_v03) and requested a download in .nc format.
#
#  Parameters for this subset
#  -----------------------------
#  Collection: IMOS - SRS - SST - L3S - Single Sensor - 6 day - day time - Australia
#  Spatial: From Lat/Lon -60.0,90.0 to Lat/Lon -5.0,180.0
#  Temporal: 2019-May-25-15:20-UTC to 2019-Jun-15-15:20-UTC
# 
# ======================================================================================
# Load SST data
```{r}
sst01<-nc_open("~/Downloads/IMOS_aggregation_20200404T015637Z/IMOS_aggregation_20200404T015637Z.nc");

sst01.d<-ncvar_get(sst01, 'sea_surface_temperature');  ### check the variable name is correct, can be 'sst4', or other names
#sst01.d[sst01.d > 42.00072]<-NA  ### beware this value should be changed to NA in the matrix
sst<-sst01.d[,,8]-273.15
sstlon <- ncvar_get(sst01, 'lon')
sstlat <- ncvar_get(sst01, 'lat')

## Arrange the data into long format for ggplot

longitude<- rep(sstlon, length(sstlat))

## This loop might take a while and would be improved by converting to an apply function
latitude  <- NA

for (i in seq_along(sstlat)){
  latitude <-c(latitude , rep(sstlat[i],length(sstlon)))
}
sst=as.vector(sst)

SSTlong<-as.data.frame(cbind(lat=latitude[1:length(sst)], lon=longitude[1:length(sst)], sst=as.vector(sst)))

```

## FINALISING THE MAP  ##
# ======================================================================================
# Create a simple regional map with Deployment locations, labels, currents and SST using ggplot.
# Repel overlapping text labels with ggrepel.
# Skip ever 2nd deployment to prevent crowding of labels
# Add axis labels, a title and export to pdf
# To do: add Marine Park boundaries
# ======================================================================================
# Generate the Map

```{r, fig.width=7, fig.height=7, fig.retina=2}

xrange <- c(105,125)
yrange <- c (-40,-8)
ggplot() + 
  geom_tile(data=SSTlong %>% filter(between(lon, 100, 125), between(lat, -42,-8)), aes(x=lon, y=lat, fill=sst)) + 
        scale_fill_gradientn(colors=oceColors9A(300), na.value = NA, name='Sea Surface Temperature (deg C)') +
  geom_polygon(data = fortify(my_spdf ), aes( x = long, y = lat, group = group), lwd=0.2, fill="#69b3a2", color="white", alpha=0.3) +
  geom_segment(data=as.data.frame(dat2), aes(x=x, xend=x2, y=y, yend=y2), lwd=0.1, arrow = arrow(type = 'open',length = unit(0.05, "cm"))) + 
  geom_point(data=in19v03.env, aes(x=Longitude, y=Latitude), color='black') +
  geom_label_repel(nudge_x=-2,segment.alpha = 0.4,label.size = 0.1, data=in19v03.env %>% 
                     distinct(Deployment, .keep_all =T) %>% 
                     filter (round(Deployment,0) %%2==0),aes(y=Latitude, x=Longitude, label=Deployment), segment.colour = 'black', color='white', fill='black') +
  labs(title='Overview and oceanographic context of microbial sampling on IN2016_v03', x= 'Longitude (deg E)', y='Latitude (deg N)') +
  geom_sf(data = world, lwd = 0) + coord_sf(xlim = xrange, ylim=yrange, expand=F)+
  theme_bw()  

  


ggsave('~/Figure_1c_Jet.pdf', height=15, width=9)

```

## MARINE PARKS SHAPE FILES  ##
# ======================================================================================
# Add Marine Parks boundaries to the map in order to highlight sampling in the Abrohlos 
# Marine Park.
# To do: add Marine Park labels
# ======================================================================================
# Generate the Map


```{r}

my_spdf <- readOGR( 
  dsn= "/Volumes/Dropbox/MarineMicrobes Dropbox/Martin Ostrowski/AustralianMarineParksEstate/Shape/" ,
  verbose=FALSE
)
ogrListLayers(dsn= "/Volumes/Dropbox/MarineMicrobes Dropbox/Martin Ostrowski/Dropbox/AustralianMarineParksEstate/Shape/")



plot(my_spdf, col="#f2f2f2", bg="skyblue", lwd=0.25, border=0 )

install.packages('broom')
library(broom)
spdf_fortified <- fortify(my_spdf %>% filter(RESNAME == 'Abrohlos'), region = "AustralianMarineParks")


my_spdf

states <- map(my_spdf, plot = FALSE, fill = TRUE)
head(states)
```


```{r}
```

