---
title: "R Notebook"
output: html_notebook
---

Authors: Martin Ostrowski, Deepa Varkey and Mark Brown
email: martin.ostrowski@mq.edu.au
Acknowledgements: This workbook likely contains mistakes. Feel free to suggest improvements to the code.
This workbook processes a selection of public and private data, kindly refrain from sharing it without consent from the contributors. 

If you would like to access to data please register your interest with [Bioplatforms Australia](http://www.bioplatforms.org.au)


#####Workbook outlining methods to extract temperature optima and ranges from Australian Microbiome zOTU data

Data is limited to the top 31 cyanobacterial zOTUs based on total abundance in the subsampled dataset

```{r}
cyano31<-read.table("~/Dropbox/zOTU_Cyano/cyano_top31_w_meta.txt", h=T, sep='\t')
```

Quick plot to observe the geographical range of sampling
```{r}

colnames(cyano31)
plot(cyano31$lon, cyano31$Lat, asp=1)

```

Remove the rows with missing data
```{r}
cyano31t<-cyano31[complete.cases(cyano31$temp),]

```
Plot the temperature density for each zOTU


```{r}
plot(density(cyano31t$temp, weights=(cyano31t[,14]/sum(cyano31t[,14]))), ylim=c(0,0.3), xlim=c(5,34), main="weighted density Thermal Niche top 31 Cyanobacterial zOTUs", xlab="Temperature ˚C")

for (i in 14:44){
lines(density(cyano31t$temp, weights=(cyano31t[,i]/sum(cyano31t[,i]))), lwd=2, col=as.character(petb.col$col[i-13]))
}
```

Describe the shape of the density curves by finding the values for one peak and the peak 1/2 height width to provide min and max values for the temperature range 
```{r}

density <- vector('list', 31)

for (i in 1:31){
  
density[[i]]<-density(cyano31t$temp, weights=(cyano31t[,(i+13)]/sum(cyano31t[,(i+13)])))
}

peak.temps <- vector('list', 31)

for (i in 1:31){
peak.temps[[i]]<-density[[i]]$x[which.max(density[[i]]$y)]
}

min.temps <- vector('list', 31)

for (i in 1:31){
min.temps[[i]]<-min(density[[i]]$x[which(density[[i]]$y>max(density[[i]]$y)/2)])
}

max.temps <- vector('list', 31)
for (i in 1:31){
max.temps[[i]]<-max(density[[i]]$x[which(density[[i]]$y>max(density[[i]]$y)/2)])
}



zotu_trange<-cbind(unlist(min.temps),unlist(peak.temps), unlist(max.temps))
colnames(zotu_trange)<-c("min", "peak","max")
row.names(zotu_trange)<-colnames(cyano31t[,14:44])

#write.table(zotu_trange, "~/Dropbox/zOTUs/MM_thermal_optima/Cyano_zOTU_Topt_range.txt", sep='\t', quote=F)
```

#####Curve fitting and plotting

n.b. the lowess model fit is redundant. All of the data required for the Topt table can be extracted from the density object x and y coordinates.

```{r}

loess.temps <- vector('list', 31)
for (i in 1:31){
loess.temps[[i]]<-loess(density[[i]]$y ~ density[[i]]$x)
}
```

Plot some examples to show how the density curve envelopes the actual data values

```{r}
par(mfrow=c(3,2))

plot(density(cyano31t$temp, weights=(cyano31t[,14]/sum(cyano31t[,14]))), ylim=c(0,0.22), xlim=c(5,34), main="weighted density Thermal Niche", xlab="Temperature ˚C")

lines(loess.temps[[1]], lty=4, lwd=4)

points(cyano31t$temp, (cyano31t[,14]/sum(cyano31t[,14])*14), cex=0.4, col=as.character(petb.col$col[1]))

plot(density(cyano31t$temp, weights=(cyano31t[,18]/sum(cyano31t[,18]))), ylim=c(0,0.22), xlim=c(5,34), main="weighted density Thermal Niche", xlab="Temperature ˚C")

lines(loess.temps[[5]], lty=4, lwd=4)

points(cyano31t$temp, (cyano31t[,18]/sum(cyano31t[,18])*12), cex=0.4, col=as.character(petb.col$col[5]))

plot(density(cyano31t$temp, weights=(cyano31t[,22]/sum(cyano31t[,22]))), ylim=c(0,0.22), xlim=c(5,34), main="weighted density Thermal Niche", xlab="Temperature ˚C")

lines(loess.temps[[9]], lty=4, lwd=4)

points(cyano31t$temp, (cyano31t[,22]/sum(cyano31t[,22])*12), cex=0.4, col=as.character(petb.col$col[9]))

plot(density(cyano31t$temp, weights=(cyano31t[,21]/sum(cyano31t[,21]))), ylim=c(0,0.22), xlim=c(5,34), main="weighted density Thermal Niche", xlab="Temperature ˚C")

lines(loess.temps[[8]], lty=4, lwd=4)

points(cyano31t$temp, (cyano31t[,21]/sum(cyano31t[,21])*12), cex=0.4, col=as.character(petb.col$col[8]))

plot(density(cyano31t$temp, weights=(cyano31t[,25]/sum(cyano31t[,25]))), ylim=c(0,0.22), xlim=c(5,34), main="weighted density Thermal Niche", xlab="Temperature ˚C")

lines(loess.temps[[12]], lty=4, lwd=4)

points(cyano31t$temp, (cyano31t[,25]/sum(cyano31t[,25])*12), cex=0.4, col=as.character(petb.col$col[12]))

plot(density(cyano31t$temp, weights=(cyano31t[,42]/sum(cyano31t[,42]))), ylim=c(0,0.22), xlim=c(5,34), main="weighted density Thermal Niche", xlab="Temperature ˚C")

lines(loess.temps[[29]], lty=4, lwd=4)

points(cyano31t$temp, (cyano31t[,42]/sum(cyano31t[,42])*12), cex=0.4, col=as.character(petb.col$col[29]))
```


Plot the density of temperature observations to show the *bias* in sampling
```{r}
plot(density(cyano31t$temp))
```

Plot all data on one plot. The pause was added for dramatic effect in order to show the overlapping distributions.

```{r}
plot(density(cyano31t$temp, weights=(cyano31t[,14]/sum(cyano31t[,14]))), ylim=c(0,0.3), xlim=c(5,34), main="weighted density Thermal Niche top 31 Cyanobacterial zOTUs", xlab="Temperature ˚C")

for (i in 1:31){
points(cyano31t$temp, cyano31t[,(i+13)]/sum(cyano31t[,(i+13)])*18, cex=0.4, col=as.character(petb.col$col[i]))
Sys.sleep(1)
}

for (i in 14:44){
lines(density(cyano31t$temp, weights=(cyano31t[,i]/sum(cyano31t[,i]))), lwd=2, col=as.character(petb.col$col[i-13]))
Sys.sleep(1)
}

for (i in 1:31){
lines(loess.temps[[i]], col=as.character(petb.col$col[i]), lty=4, lwd=5)
}



```

#### To Do

unrelated code below implementing an uncontrolled gams model

```{r}
#colnames(cyano31t)


mod.full<-gam(cyano31t[,18] ~ temp + nox +sal + sil + po4 + DayLength +Depth +N.P, family=poisson(link="log"), data= cyano31t)

summary(mod.full)
```
```{r}

par(mfrow=c(2,4))
gam::plot.Gam(mod.full, se=T)
```

